---
title: "Avocado Prices Case Study"
author: "Leopoldine Mirtil"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Source

This analysis uses the [Avocado Prices](https://www.kaggle.com/datasets/neuromusic/avocado-prices) data set which was made publicly available by Justin Kiggins on Kaggle.com. It contains historical data about avocado prices and sales throughout the United States. 

 
### Objective

The following tasks: 

1. Identify the cities that had the cheapest and most expensive prices per year 
2. Determine the average price and sales by type 
3. Find the top 10 cities with the overall lowest and highest prices
4. Analyze the relationship between prices and quantity


## Import Data

### Step 1 - Import Data

#### Load Library

```{r, message=FALSE, warning=FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
library(lubridate)
library(plotly)
```

#### Set Directory & Import Data

```{r}
df <- read.csv('avocado.csv')
```


### Step 2 - Clean Data

```{r}
#drop index column
df <- df[-c(1)]

#change data type of date 
df$Date <- as.Date(df$Date)

#add Col
df$Sales = df$AveragePrice * df$Total.Volume

#rename cols 
df <- rename(df, TotalQuantity=Total.Volume, TotalBags=Total.Bags, SmallBags=Small.Bags, LargeBags=Large.Bags, 
             XLargeBags=XLarge.Bags, Type=type, Year=year, Region=region)

#remove PLU cols 
df <- df[-c(4:6)] 
```


### Step 3 - Inspect Data

#### Check for Missing/Null Values

```{r}
sapply(df, function(x) sum(is.na(x)))
```

#### Check for Empty Values

```{r}
sapply(df, function(x) sum(x==""))
```

#### Check Date Ranges per Year

```{r}
df %>%
  group_by(Year) %>%
  reframe(DateRange = range(Date)) %>%
  arrange(Year)
```

#### Remove Rows

```{r}
#remove 2018 rows
df <- df[!(df$Year %in% 2018),]

#confirm removal
unique(df$Year)
```


#### View Modified Dataframe

```{r}
str(df)
```


```{r, echo=FALSE}
#export modified file
write.csv(df, 'avocadoPrices.csv', row.names = FALSE)
```


### Step 4 - Analyze Data

#### Descriptive Analysis

```{r}
summary(df)
```

#### Summary Statistics of Region Prices

```{r}
df %>%
  group_by(Region) %>%
  summarise(MinPrice=min(AveragePrice),
            MeanPrice=round(mean(AveragePrice),2),
            MedianPrice = round(median(AveragePrice), 2),
            MaxPrice=max(AveragePrice)) %>%
  arrange(MinPrice) 
```


#### Annual Minimum & Maximum Prices 

```{r, warning=FALSE, message=FALSE}
df %>%
  group_by(Year, Region) %>%
  summarise(PriceRange = range(AveragePrice), .groups="drop") %>%
  group_by(Year) %>%
  filter(PriceRange == range(PriceRange))
```


#### Top 10 Regions by Cheapest Price

```{r}
df %>%
aggregate(cbind(AveragePrice, TotalQuantity, Sales) ~ Region, min) %>%
rename (MinPrice = AveragePrice,
        Quantity = TotalQuantity) %>%
arrange(MinPrice) %>%
slice(1:10)
```


#### Top 5 Regions with Frequently Cheap Prices Per Year

```{r}
#create mode function
mode_func <- function(x) {
  u <- unique(x)
  tab <- tabulate(match(x, u))
  u[tab == max(tab)]
}

df %>%
  group_by(Year, Region) %>%
  reframe(CommonPrice= mode_func(AveragePrice)) %>%
  arrange(Year, Region, CommonPrice) %>%
  group_by(Year, Region) %>%
  slice_head(n=1) %>%
  arrange(Year, CommonPrice) %>%
  group_by(Year) %>%
  slice(1:5) 
```

#### Top and Bottom 10 Regions by Median Prices 

```{r}
df %>%
  group_by(Region) %>%
  reframe(MedianPrice= round(median(AveragePrice), 2)) %>%
  arrange(MedianPrice) %>%
  slice(unique(c(1:10, n() - 9:0))) 
```

#### Average Price and Sales by Type

```{r}
df %>%
  group_by(Type) %>%
  summarise(AvgPrice = round(mean(AveragePrice),2), 
            AverageSales = mean(Sales)) %>%
  arrange(AvgPrice)
```


### Step 5 - Visualizations

```{r, echo=FALSE, warning=FALSE}
df %>%
  group_by(Region) %>%
  summarise(MedianPrice= round(median(AveragePrice), 2)) %>%
  arrange(MedianPrice) %>%
  slice(unique(c(1:10, n() - 9:0))) %>%
  ggplot(aes(y=reorder(Region, -MedianPrice), x=MedianPrice, fill=ifelse(MedianPrice < 1.50, "darkgreen", "darkred"))) +
  geom_col() +
  ggtitle(paste("Top & Bottom 10 Lowest Avocado-Priced Regions")) +
  scale_fill_identity() +
  theme(legend.position = "none", axis.title.y=element_blank(), axis.title.x=element_blank()) 
```


```{r, echo=FALSE, warning=FALSE}
df %>%
  group_by(Year, Region) %>%
  reframe(Prices = range(AveragePrice)) %>%
  group_by(Year) %>%
  filter(Prices == range(Prices)) %>%
  mutate(Cost = ifelse(Prices < 1, "Cheapest", "Expensive")) %>%
  plot_ly(x = ~Year, y = ~Prices,  color=~Cost, type = 'bar', customdata=~Cost, text=~Region, 
          hovertemplate=paste(' <b>%{customdata}</b><br>', 
                              '%{x}: %{text}, $%{y:.2f}<br><extra></extra>')) %>%
  layout(title="Annual Lowest and Highest Avocado Prices")
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
df %>%
  group_by(Region) %>%
  summarise(Price= round(median(AveragePrice), 2),
            Quantity=median(TotalQuantity), .groups="drop") %>% 
  arrange(Price) %>%
  plot_ly(type="scatter", mode="markers", y = ~Quantity, x = ~Price, size=~Price, text= ~Region,
          hovertemplate=paste(' <b>%{text}</b><br>', 
                          'Price: $%{x:.2f}<br>',
                          'Quantity: %{y:,.2f}<br><extra></extra>')) %>%
  layout(title="2015-2017 Median Avocado Prices and Quantity")
```
